---
title: "Midterm"
author: "Cameron Kay & Ashley Miller"
date: "4/22/2019"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include = FALSE}
# load required packages
library(tidyverse)
library(glue)
library(rio)
library(pracma)
library(magrittr)
```

## Part A: Data

### Question 1 (5 points) 

```{r parta_question1_example}
download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}
```

Use the function above to download all the data for each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

Conduct some basic data cleaning to make your data file look like the following.

  + Filter for only student groups coded as "White" or "Hispanic/Latino".
  + Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.
  + Remove any row that has missing data in any of the n variables
  
```{r parta_question1_cam, cache = TRUE}
# download the data
data <- map_df(15:18, download_file)

# tidy the data
data %<>%
  # clean the column names
  janitor::clean_names() %>%
  # filter for only white and hispanic/latino students
  filter(student_group == "White" |
         student_group == "Hispanic/Latino") %>%
  # drop percentages or collapsed levels 
  select(academic_year,
         district,
         school,
         student_group,
         grade_level,
         starts_with("number_level")) %>%
  # gather by level
  gather(key = "level", value = "n", starts_with("number_level")) %>%
  # parse level number
  mutate(level = parse_number(level)) %>%
  # drop rows with missing data
  filter(!is.na(n)) %>%
  # arrange in same order as midterm example (for sanity)
  arrange(academic_year,
          district,
          student_group,
          grade_level,
          level)
  
```

### Question 2 (2 points)

Calculate the cumulative n for each school by student group, grade, and academic year. The result should look like the below. Hint, look at `?base::cumsum.`

```{r parta_question2_cam}
# create a cumulative n column
data %<>%
  group_by(student_group, grade_level, academic_year) %>%
  mutate(cn = cumsum(n))
```

### Question 3 (3 points)

Reformat the data so it looks like the below, removing `n` and filling by `cn`. Remove rows that have missing data for either student group.

```{r parta_question3_cam}
# reformat data
data %<>%
  # drop n
  select(-n) %>%
  # spread cn by student group
  spread(key = student_group, value = cn) %>%
  # clean new column names
  janitor::clean_names() %>%
  # drop rows with missing cn
  filter(!is.na(hispanic_latino),
         !is.na(white))
```

## Part B: Achievement gaps

The function below estimates the average difference between two distributions in terms of an effect size. In this case, we are using the cumulative counts to approximate the empirical cumulative distribution function for each group. The distance between the distributions is then estimated and transformed to an effect size-like measure (for more information, see Ho & Reardon, 2012). The nice thing about this approach, is that we’re able to obtain an effect size on the average difference in achievement between to groups of students as if we had the full, student level data even though we just have the counts within each category.

In the below function, the first argument supplied is the data source, followed by two string variables, the names of the reference and focal distributions, respectively (e.g., "white" and "hispanic_latino" in this case).

Note - you’ll need to install the {pracma} package first (assuming you don’t have it installed already, of course).

### Question 1 (10 points) 

```{r parta_question1_example}
# create function
gap <- function(data, ref, foc) {
    x <- data[[ref]]
    y <- data[[foc]]
    auc <- pracma::trapz(y / y[length(x)],
                         x / x[length(x)])
    sqrt(2) * qnorm(auc)
}

# example
gap(data[1:4, ], "white", "hispanic_latino")
```

Estimate an achievement gap effect size for every school in the state that reported data on both student groups (i.e., using the data we created above), for each grade level in each academic year.

```{r partb_question1_cam}
# create a unique identifier using school, academic year, and grade level
data_es <- data %>%
  # combine district, school, grade_level, and academic_year
  mutate(id = paste(district, school, grade_level, academic_year)) %>%
  # replace spaces with underscores and make the entire id lowercase
  mutate(id = gsub(" ", "_", id),
         id = tolower(id))

# split using the id column
data_es_split <- data_es %>%
  split(data_es$id)

# calculate the effect size for each unique id and save as a vector of doubles
es <- map_dbl(data_es_split, gap, ref = "white", foc = "hispanic_latino")

# prepare dataframe for addition of effect sizes
data_es %<>%
  # retain only one column for each school/year/grade combination
  filter(level == 1) %>%
  # arrange using the unique id
  arrange(id) %>%
  # drop the id and level columns
  select(-id, -level)

# add effect sizes to the dataframe
data_es$effect_sizes <- es
```

### Question 2 (10 points) 

The plot below shows the achievement gap estimates for one school by grade in Ashland School District during the 2017-18 school year. Produce a similar plot to the below (noting the school, academic year, and school district) for each of the first 100 unique school/year/district combinations. Hint - you’ll want to use your effect size data from from the previous question, nest it, then apply `slice(1:100)`. Note that the only reason I’m asking you to slice the data frame is just to reduce run time. In reality, you would do this for all school/year/district combinations.

```{r partb_question2_ash}
# use the data called data_es

```

### Question 3 (10 points) 

Save the plots into a “plots” directory. Make sure the file names are meaningful.

```{r partb_question3_ash}

```
